Introduction: \citet{kuriwaki2026roleconfounderslinearityecological}, but also \citet{mccartan2025identificationsemiparametricestimationconditional}.


I should say something about the more general importance of estimating correctly when we do not observe individual-level data.

I should say that sometimes we do not observe individual-level data and we might be tempted to infere something based on highly granular data. But these inferences may be wrong.

That could be relevant for economists in various applications and analyses and it is not solely a nuisance for econometricians.

In fact, data are usually reported by official bureau of statistics even at highly granular level but not at individual level and some relationships become masked.


Subsequently, I should present the problem with appropriate notation.

And probably I should also mark the difference between this problem and others problem of aggregation such as the one discussed in BLP model.

In any case, I will later decide how to structure the dissertation and where to put the complete explanation of the method: in principle, I would put it into the introduction and later I would present the syntethic example just referring to what I have already written here.


%In particolare, quali sono le differenze fra questo metodo e gli altri?

%Rispetto agli effetti fissi, qua dico che la variabilità non è una semplice trasposizione data dagli effetti fissi (dico ad esempio che tra Pennsylvania e New Mexico in New Mexico i laureati votano il 10\% più Democratico, ma non perché il New Mexico sia spostato tutto 10\% più Democratico; magari lo sono i laureati, mentre i non laureati sono il 5\% meno Democratici rispetto a quelli della Pennsylvania).

%Invece Gary King proponeva di fare i coefficienti Beta con una certa variabilità nelle diverse geografie ma secondo una distribuzione parametrica (Normale).

%E così via, mostrando un po' di quello che mostrano in \citet{kuriwaki2026roleconfounderslinearityecological}.

%Un'altra possibilità, direi, è quella di fare una regressione usando tanti regressori, dicendo che quello precedente non è un problema di ecological fallacy ma di omitted variables.

%Posso dire che l'altra differenza è quella con l'approccio di modellare le interazioni tra agenti che è molto più complicato.
